Species richness and network structure jointly drive total biomass and its temporal stability in fish communities

### 2020-12-07 manuscript/bef_stability/manuscript.Rmd
Discussion reviewed according to Elisa

Overall, its study is unique by its level of integration, from environmental
effects to total biomass and stability, going through species richness and
network structure. It is the first direct assessment of a possible joint effect
of species richness and network structure on both total biomass and its temporal
stability. We found that species richness and mean trophic level had a positive
effect on total biomass. In contrast, species richness had a negative effect on
temporal stability whereas mean trophic level covaried positively. We also found
that the effect of species richness and mean trophic level on total biomass and
its temporal stability were not independant as species richness had a negative
effect on mean trophic level.  Furthermore, we found that connectance had no
significant effect on both total biomass and its temporal stability. Finally, we
found that the magnitude of the effects of average temperature and altitude was
equal or superior to the effects of species richness and mean trophic level.

@@

We found that species richness increased total biomass as found in salt marshes,
grasslands and forests [@hector_plant_1999; @tilman_biodiversity_2014;
@wu_relationship_2015; @li_relationship_2018] as well as in streams
[@woods_testing_2020]. The positive effect of species
richness on total biomass might be due to resource use complementarity
[@carey_determining_2011], which is held by niche differentiation
[@loreau_biodiversity_1998; @loreau_partitioning_2001] or to the higher
probability of getting high biomass species with higher species richness
[@loreau_biodiversity_2001].

Surprisingly, we found a negative relationship between temporal stability and
species richness. Overall, it contrasts with experimental studies in grasslands
[@tilman_biodiversity_2006; @valencia_synchrony_2020]; mesocosms (Pennekamp et al. 2018), in 
natural communities of butterflies [@olivier_urbanization_2020] or even in lotic
fish communities [@franssen_annual_2011] where positive relationships were found
between temporal stability and species richness. However, other studies found an
opposite relationship in natural grassland [@yang_effects_2011] and a marginal
one in natural forests [@declerck_species_2006].

@@

The positive effect of species richness on
$\overline{CV_{sp}}$, thereby decreasing stability, is more intriguing. While
positive effect of species richness on $\overline{CV_{sp}}$ was reported in
experimental grasslands and natural bird communities [@tilman_biodiversity_2006;
@olivier_urbanization_2020], negative and neutral relationships have been found
in butterfly and bat communities. Our results contrast with a previous
meta-analysis which concludes that species richness had a positive effect on
population stability in multi-trophic aquatic microcosms
[@jiang_different_2009]. In foodwebs, the relationship between temporal
stability of populations and species richness depends on interaction strength
and its distribution [@mccann_reevaluating_1997; @thebault_trophic_2005;
@brose_allometric_2006], high interaction strength inducing a negative
relationship. Therefore, it is possible that our results are due to a high
degree of omnivory in foodwebs, which combined with strong trophic interactions
lead to more fluctuations in population biomass.

@@

## Average trophic level increased total biomass and its temporal stability

The average trophic level was positively linked to total biomass.
Experimental and theoretical work suggested that a higher trophic level can be the
result of a stronger top-down effect, meaning higher predation rates and biomass
fluxes to higher trophic levels [@kratina_warming_2012], hence leading to an
increase in total biomass with trophic length. Furthermore,
@wang_biodiversity_2016 have shown that total nutrient uptake might increase
with the length of trophic chains and so results in higher total biomass.

In contrast, connectance had no effect on total biomass. One might have expected
that an increase in connectance may translate in a higher total nutrient uptake
and so results in higher total biomass in the fish compartment. A reason for
such an absence of an effect might be explained by the fact that our data lack
information about the lower compartments of the community, i.e., the primary
producers, zooplankton and zoobenthos.

Mean trophic level had a positive effect on biomass stability by its negative
effect on $\overline{CV_{sp}}$, thereby increasing average species stability. It
might result from a higher stability of higher trophic level as predicted by
food chain models in the presence of strong top-down control
[@barbier_pyramids_2019; @shanafelt_stability_2018]. But it
might not hold when considering omnivorous species
[@wang_biodiversity_2016]. An alternative explanation is that higher mean
trophic level leads to a stronger top-down control and resource complementarity
and thus reduce biomass fluctuations in lower trophic levels
[@rooney_structural_2006].

We found that connectance had no effect on temporal stability. This result
contradicts previous precedent theoretical studies on the relationship between
stability and network structure [@angelis_stability_1975; @may_will_1972;
@thebault_stability_2010; but see @neutel_reconciling_2007], although they
considered different stability concepts and metrics [@arnoldi_resilience_2016].
However, our measure of connectance does not include the variability of the
lower compartments of the foodweb and further research should examine the effect
of connectance on temporal stability in more complete empirical foodwebs.

@@

In contrast with previous studies, we found that the average population
variability ($\overline{CV_{sp}}$) had a stronger negative effect on biomass
stability than synchrony [@tilman_biodiversity_2006;
@olivier_urbanization_2020]. It explains why we observe a negative effect of
species richness on temporal stability of biomass. The lower effect of synchrony
might be explained by the presence of generalist predators in stream
communities, that raise synchrony between prey species that share the
same predator and thus reduces compensatory
population dynamics [@raimondo_interspecific_2004]. Alternatively, it is
possible that, in aquatic foodwebs, the occurrence of large oscillatory dynamics
 leads to a higher contribution of population
variability rather than synchrony to community stability
[@mccann_reevaluating_1997]. How both strengths and the relative contribution of
synchrony and population variaribility to overall stability vary between
foodwebs and single trophic communities is still an open question.

@@

We found that the total effect of environmental drivers on total biomass and its
stability was indirect and of the same magnitude order than the effect of
species richness and average trophic level, highlighting that environmental
changes may result in signicant ecosystem responses, through alteration of
both species richness and network structure.

Consistent with research showing that foodwebs in wider streams support more
species and biomass [@allan_stream_2007], we found that the PCA axis related to
stream size had a positive effect on total
biomass through a positive effect on mean trophic level. Greater ecosystem size
might provide a greater diversity of resources through habitat heterogeneity,
and so higher species richness and total biomass by surface unit, through
resource complementarity. Greater diversity of resources can also maintain
more biomass at high trophic level [@post_ecosystem_2000; @doi_resource_2009].
Furthermore, the positive effect of the PCA axis related to temperature on total
biomass was of the magnitude of effects of network structure and species
richness, in line with the mounting body of work reporting important effects of
temperature on the structure and biomass of aquatic system [e.g.
@dossena_warming_2012; @edeline_ecological_2013; @emmrich_geographical_2014;
@hattab_forecasting_2016; @woods_testing_2020]. Contrasting with previous studies in experimental
plant communities [@grace_integrative_2016], aquatic mesocosms [@yvondurocher_warming_2011; @dossena_warming_2012] and marine fish communities [@maureaud_biodiversityecosystem_2019]
, we found that environmental drivers overall affect total biomass indirectly
rather than directly [@woods_testing_2020].

>>>>

Overall, this study is unique by its level of integration, from environmental
effects to total biomass and stability, going through species richness and
foodweb structure. It is the first direct assessment of a possible joint effect
of species richness and network structure on both total biomass and its temporal
stability. We found that species richness and mean trophic level had a positive
effect on total biomass. In contrast, species richness had a negative effect on
temporal stability whereas mean trophic level covaried positively. We also found
that the effect of species richness and mean trophic level on total biomass and
its temporal stability were not independent as species richness had a negative
effect on mean trophic level. Furthermore, we found that connectance had no
significant effect on both total biomass and its temporal stability. Finally, we
found that the magnitude of the effects of average temperature and altitude was
equal or superior to the effects of species richness and mean trophic level.

@@

We found that species richness increased total biomass as already found in salt marshes,
grasslands and forests [@hector_plant_1999; @tilman_biodiversity_2014;
@wu_relationship_2015; @li_relationship_2018] as well as in streams
[@woods_testing_2020]. The positive effect of species
richness on total biomass might be due to resource use complementarity
[@carey_determining_2011], which is held by niche differentiation
[@loreau_biodiversity_1998; @loreau_partitioning_2001] or to the higher
probability of getting high biomass species with higher species richness
[@loreau_biodiversity_2001]. Our result is also in agreement with a recent
study predicting a positive relationship between animal diversity and total
consumer biomass in size-structured food web models [@schneider_animal_2016].

Surprisingly, we found a negative relationship between temporal stability and
species richness. Overall, it contrasts with experimental studies in grasslands
[@tilman_biodiversity_2006; @valencia_synchrony_2020], mesocosms
[@pennekamp_biodiversity_2018] with studies on natural communities such as
butterflies  or lotic fishes [@olivier_urbanization_2020; @franssen_annual_2011;
@houlalan__2018], where positive relationships were found between temporal
stability and species richness. However, some other studies also found a
negative relationship, for instance in natural grasslands [@yang_effects_2011]
or in natural forests as a marginal trend [@declerck_species_2006].

@@

The positive effect of species richness on
$\overline{CV_{sp}}$, thereby decreasing stability, is more intriguing. While
studies focusing on single-trophic communities have reported mixed effects of
species richness on $\overline{CV_{sp}}$, with destabilising effects of species
richness at population level in grasslands [@jiang_different_2009;
@gross__2014], a previous meta-analysis suggested that species richness had
often a positive effect on population stability in multi-trophic communities
[@jiang_different_2009].
In foodwebs, the relationship between temporal
stability of populations and species richness depends on interaction strength
and its distribution [@mccann_reevaluating_1997; @thebault_trophic_2005;
@brose_allometric_2006 DOWNING à la place de BROSE?], high interaction strength inducing a negative
relationship whereas the presence of weak interactions at high species richness is
expected to promote stability. Therefore, it is possible that our results are
due to a high degree of omnivory in foodwebs, which combined with strong trophic
interactions lead to more fluctuations in population biomass.

@@

## Average trophic level increased total biomass and its temporal stability while connectance had no effect

The average trophic level was positively linked to total biomass.  Furthermore,
@wang_biodiversity_2018 have shown that total nutrient uptake might increase
with the length of trophic chains and so results in higher total biomass.  While
the predictions of @wang_biodiversity_2018 were restricted to food chains, and
thus might not hold in more complex food webs with omnivorous species, our
results are in agreement with their model predictions and highlight the
importance of food chain length as a determinant of total biomass in foodwebs.
In addition, mean trophic level had a positive effect on biomass stability by
its negative effect on $\overline{CV_{sp}}$, thereby increasing average species
stability. It might result from a higher stability of higher trophic level as
predicted by food chain models in the presence of strong top-down control
[@barbier_pyramids_2019; @shanafelt_stability_2018]. An alternative
explanation is that higher mean trophic level leads to a stronger top-down
control and resource complementarity and thus reduce biomass fluctuations in
lower trophic levels [REF?].
In contrast, connectance had no effect on total biomass. One might have expected
that an increase in connectance may translate in a higher total nutrient uptake
and so results in higher total biomass in the fish compartment. 
However, a few studies suggest that greater connectance should decrease
complementarity in resource use between species, which is expected to lead to
lower total biomass [@thebault_food-web_2003; @poisot_trophic_2013].
A reason for
such an absence of an effect might be explained by the fact that our data lack
information about the lower compartments of the community, i.e., the primary
producers, zooplankton and zoobenthos.

Regarding temporal stability, our results
contradicts numerous previous theoretical studies on the relationship between
connectance and stability [e.g., @angelis_stability_1975; @may_will_1972;
@thebault_stability_2010; but see @neutel_reconciling_2007], although they
considered different stability concepts and metrics [@arnoldi_resilience_2016].
The lack of effect of connectance on the temporal stability of total biomass
might be related to the fact that our measure of connectance does not take into
account the distribution of interaction strengths in the food webs. Indeed, food
web connectance is expected to have different consequences on the temporal
stability of total biomass depending on the strength of trophic interactions
[@thebault_trophic_2005]. However, our measure of connectance does not fully
account for the whole foodweb structure, and further research should examine the
effects of connectance on total biomass and its temporal stability in more
complete empirical foodwebs.

@@

In contrast with previous studies, we found that the average population
variability ($\overline{CV_{sp}}$) had a stronger negative effect on biomass
stability than synchrony [@tilman_biodiversity_2006;
@olivier_urbanization_2020]. It explains why we observe a negative effect of
species richness on temporal stability of biomass. The lower effect of synchrony
might be explained by the presence of generalist predators in stream
communities, that raise synchrony between prey species that share the
same predator and thus reduces compensatory
population dynamics [@raimondo_interspecific_2004]. Alternatively, it is
possible that, in aquatic foodwebs, the occurrence of large oscillatory dynamics
 leads to a higher contribution of population
variability rather than synchrony to community stability
[@mccann_reevaluating_1997]. How both strengths and the relative contribution of
synchrony and population variability to overall stability vary between
foodwebs and single trophic communities is still an open question.

@@

We found that the total effect of environmental drivers on total biomass and its
stability was mainly indirect and of the same magnitude order than the effect of
species richness and average trophic level, highlighting that environmental
changes may result in significant ecosystem responses, through alteration of
both species richness and network structure.

Contrasting with previous studies in experimental plant communities
[@grace_integrative_2016], aquatic mesocosms [@yvondurocher_warming_2011;
@dossena_warming_2012] and marine fish communities
[@maureaud_biodiversityecosystem_2019], we found that environmental drivers
overall affect total biomass indirectly rather than directly
[@woods_testing_2020]. First, consistent with research showing that foodwebs in
wider streams support more species and biomass [@allan_stream_2007], we found
that the PCA axis related to stream size had a positive effect on total biomass
through a positive effect on mean trophic level and species richness. Greater
ecosystem size might provide a greater diversity of resources through habitat
heterogeneity, and so higher species richness and total biomass by surface unit,
through resource complementarity. Greater diversity of resources can also
maintain more biomass at high trophic level [@post_ecosystem_2000;
@doi_resource_2009].  Second, the positive effect of the PCA axis related to
temperature on total biomass was of the magnitude of the effects of network
structure and species richness, in line with the mounting body of work reporting
important effects of temperature on the structure and biomass of aquatic systems
[e.g.  @dossena_warming_2012; @edeline_ecological_2013;
@emmrich_geographical_2014; @hattab_forecasting_2016; @woods_testing_2020].  We
found that streams with higher average temperature supported higher diversity as
well as higher average trophic level and total biomass, which is in agreement
with recent findings on streams where the positive effects of temperature were
related to larger nutrient supply [@ogorman_unexpected_2017].

While the PCA axis related with stream size had only a very weak effect on the
temporal stability of total biomass, the PCA axis related to temperature had a
clear negative effect on the temporal stability through a direct destabilizing
effect on populations and through indirect effects. Temperature can have both
stabilizing and destabilizing effects in resource-consumer models depending on
the temperature optima of species growth rates and consumption rates and on how
nutrient supply varies with temperature (Uszko et al. 2017). In our case, the
observed increased variability of populations might also relate with factors
covarying with temperature such as the average Biological Organic Demand which
is linked to enrichment.
